Machine Learning for Handwriting Recognition: Algorithms and Techniques

### 2023-11-12 Manuscript/Manuscript.tex
General Commits - Adding Contents

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage[margin = 1in]{geometry}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{natbib}
\usepackage{lipsum}
\usepackage{setspace}
\usepackage[colorlinks=true, citecolor=blue]{hyperref}
\doublespacing
\title{Machine Learning for Handwriting Recognition: Algorithms and Techniques}
\author{Lai Jiang\\
  University of Connecticut Department of Statistics
}

@@

\section* {Data Collection}
When dealing with handwriting recognition tasks, choosing the right libraries and packages is crucial to obtain reliable training results. Notable libraries include TensorFlow and Keras, which provide a wealth of deep learning capabilities that are essential for building and training neural networks that can recognize handwritten characters. Additionally, PyTorch is a highly flexible and dynamic open source machine learning framework, particularly known for its easy of using and suitability for a wide range of tasks, including handwriting recognition. It can build, train, and evaluate deep neural networks through a simple Python interface. On the other hand, the MNIST dataset is a classic dataset in the machine learning community, containing 60,000 training images of handwritten digits 0 to 9 and 10,000 test images. Each image is represented in grayscale and is 28 x 28 pixels in size. This dataset has been widely used as a benchmark to evaluate and compare the performance of different algorithms in the field of handwriting recognition. When used in conjunction with PyTorch, the MNIST dataset serves as a powerful tool for efficiently developing and evaluating handwriting recognition models.

@@

Data Augmentation contains a subset, called the Color space augmentations. It can be used as adjust the brightness or contrast. These methods are particularly useful for colored ink or paper. Elastic distortions, which warp the image in a non-linear manner, effectively mimic the natural irregularities in handwriting. Further, synthetic data generation, where entirely new handwriting samples are generated using advanced techniques like Generative Adversarial Networks (GANs), offers a way to significantly expand the dataset, especially in scenarios where data collection is challenging or privacy concerns are paramount. These augmentation techniques not only help in creating a more robust model by providing a wide range of handwriting styles and conditions but also aid in reducing overfitting, ensuring that the model generalizes well to new, unseen data. The strategic use of data augmentation in handwriting recognition thus plays a crucial role in enhancing the accuracy and reliability of the resulting machine learning models.

\section* {Algorithms in Handwriting Recognition}
However, collecting and augmenting the data cannot do everything for us, As I mentioned previously, the MNIST dataset contains 60000 for training purposes and 10000 for optimizing the model, we have to choose a relatively credible approach to recognize what the user have written.

\bibliographystyle{plain}
\bibliography{Citations.bib}
\cite{824821}
\cite{NIPS2008_66368270}
\cite{6981034}
\end{document}

>>>>

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage[margin = 1in]{geometry}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{natbib}
\usepackage{lipsum}
\usepackage{setspace}
\usepackage{tikz}
\usepackage[colorlinks=true, citecolor=blue]{hyperref}
\doublespacing
\title{Machine Learning for Handwriting Recognition: Algorithms and Techniques}
\author{Lai Jiang\\
  University of Connecticut Department of Statistics
}

@@

\section* {Data Collection}
When we're dealing with handwriting recognition tasks, choosing the right libraries and packages is crucial to obtain reliable training results. Notable libraries include TensorFlow and Keras, which provide a wealth of deep learning capabilities that are essential for building and training neural networks that can recognize handwritten characters. Additionally, PyTorch is a highly flexible and dynamic open source machine learning framework, particularly known for its easy of using and suitability for a wide range of tasks, including handwriting recognition. It can build, train, and evaluate deep neural networks through a simple Python interface. On the other hand, the MNIST dataset is a classic dataset in the machine learning community, containing 60,000 training images of handwritten digits 0 to 9 and 10,000 test images. Each image is represented in grayscale and is 28 x 28 pixels in size. This dataset has been widely used as a benchmark to evaluate and compare the performance of different algorithms in the field of handwriting recognition. When used in conjunction with PyTorch, the MNIST dataset serves as a powerful tool for efficiently developing and evaluating handwriting recognition models.

@@

Data Augmentation also contains a subset, called the Color space augmentations. It can be used as adjust the brightness or contrast. These methods are particularly useful for colored ink or paper. The Elastic distortions would warp the image in a non-linear manner, effectively mimic the natural irregularities in handwriting. Further, synthetic data generation, where entirely new handwriting samples are generated using advanced techniques like Generative Adversarial Networks (GANs), offers a way to significantly expand the dataset, especially in scenarios where data collection is challenging or privacy concerns are paramount. These augmentation techniques not only help in creating a more robust model by providing a wide range of handwriting styles and conditions but also aid in reducing overfitting, ensuring that the model generalizes well to new, unseen data. The strategic use of data augmentation in handwriting recognition thus plays a crucial role in enhancing the accuracy and reliability of the resulting machine learning models.

\section* {Algorithms in Handwriting Recognition}
However, collecting and augmenting the data cannot do everything for us, As I mentioned previously, the MNIST dataset contains 60000 for training purposes and 10000 for optimizing the model, we have to choose a relatively credible approach to recognize what the user have written. For approaches, we now have: Traditional Machine Learning Approaches, Neural Network Approaches and Hybrid. Each of them contains various methods and concepts.

\paragraph{Traditional Machine Learning Approaches}
The Traditional Machine Learning Approaches is the most fundamental and a little bit of antiquated way. The most well-known of these is the decision tree. The decision tree belongs to the supervised learning and it can tackle classification and regressions tasks. The process of it looks like a flowchart, and each node on the chart contains the "test result" of the event. Decision trees provide an intuitive way to extract decision rules from data. It can be visualized and is easy to understand, even for people without a technical background. At the same time, it also reveals the structure and patterns in the data, helping us understand how decisions are supported by data. For example, Suppose we have a data set about weather conditions and whether outdoor activities are performed, we can setup several parameters to estimate it, such as "Weather", which contains "Sunny, Cloudy, and Raining"; "Temperature", which contains "Hot, Proper, Cold", and the third parameter called "Do we want to do the outdoor activities" which only includes yes or no. Our goal is to create a decision tree that helps us decide whether to engage in outdoor activities based on weather conditions. On the climax of the tree, is "Do we want to do the outdoor activities", then, in the first layer of the tree, are the parameters of the weather "Sunny, Cloudy, and Windy" And after this layer, we can output the decisions with yes or no. In this case, sunny and cloudy, could be "yes", and "no" for the rainy. And after many layers of iteration, it could finally have the ability of predictions.

Other approaches of Traditional Machine Learning includes

\paragraph{Neural Network Approaches}
Although there are many ways to emphasize and implement the machine learning and training process, the most common way that related to neural network could be elaborated like this: For example, if we have a hand-written number, say 7, in this case, how should we let the machine know it is 7? In the MNIST Dataset, all pictures in it are 5 to 5 scaled, which means there are 25 pixels in total. And then we can line them up into a one-dimensional column, and label them 1 if the block or pixel is

\section{Further }
\bibliographystyle{plain}
\bibliography{Citations.bib}
\cite{824821}
\cite{NIPS2008_66368270}
\cite{6981034}
\end{document}
